# Database creation {#creation}


```{r start, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Start

Install and open packages `RSQLite`

```{r install, eval = FALSE, echo = TRUE}
install.packages("RSQLite")
library(RSQLite)
```
 
Establish database connection

Connecting to the term_project database

```{r db_connect, eval = FALSE, echo = TRUE}
term_project <- dbConnect(drv = RSQLite::SQLite(),
                        "C:/Users/A02345823/Box/Lab_Group/Kipling/Classes/WILD6900_EcoRepSci/Term_project/Term_project/database/severity.db")
```

Create severity table
Term project database creation based on structure in figure 1

```{r severity, eval = FALSE, echo = TRUE}
dbExecute(term_project, "CREATE TABLE severity(
severity_id integer NOT NULL PRIMARY KEY,
MTBSdNBR float,
severity float
);")
```

Create forest behavior table

```{r forest_behavior, eval = FALSE, echo = TRUE}
dbExecute(term_project, "CREATE TABLE forest_behavior(
forest_behave_id integer NOT NULL PRIMARY KEY,
severity_id float,
flmlnA float,
flmlnE float,
flmlnH float,
flmlnL float,
flmlnM float,
ftype_E float,
ftype_H float,
ftype_L float,
ftype_M float,
ros float,
rosE float,
rosH float,
rosL float,
rosM float,
FOREIGN KEY(severity_id) REFERENCES severity(severity_id)
);")
```

Create forest risk table

```{r forest_risk, eval = FALSE, echo = TRUE}
dbExecute(term_project, "CREATE TABLE forest_risk(
forest_risk_id integer NOT NULL PRIMARY KEY,
severity_id float,
fri float,
fti float,
sdr varchar(12),
WHP float,
FOREIGN KEY(severity_id) REFERENCES severity(severity_id)
);")
```

Create forest_structure table

```{r forst_structure, eval = FALSE, echo = TRUE}
dbExecute(term_project, "CREATE TABLE forest_structure(
str_id integer NOT NULL PRIMARY KEY,
severity_id float,
wwacbd float,
wwacbh float,
wwacc float,
wwach float,
wwafuel13 float,
wwafuel40 float,
FOREIGN KEY(severity_id) REFERENCES severity(severity_id)
);")
```

Create forest_topo table

```{r forest_topo, eval = FALSE, echo = TRUE}
dbExecute(term_project, "CREATE TABLE forest_topo(
forest_topo_id integer NOT NULL PRIMARY KEY,
severity_id float,
asp varchar(5),
dem float,
slope float,
FOREIGN KEY(severity_id) REFERENCES severity(severity_id)
);")
```


Create forest_veg table

```{r forest_veg, eval = FALSE, echo = TRUE}
dbExecute(term_project, "CREATE TABLE forest_veg(
forest_veg_id integer NOT NULL PRIMARY KEY,
severity_id float,
EVT varchar(60),
SystGrp varchar(60),
SAFSrm varchar(60),
NVCSOrd varchar(16),
NVCSclass varchar(25),
NVCSsubclass varchar(40),
SYSTMGRPPH varchar(30),
FOREIGN KEY(severity_id) REFERENCES severity(severity_id)
);")
```

Create non-forest db tables

Create non-forest behavior table

```{r nf_behavior, eval = FALSE, echo = TRUE}
dbExecute(term_project, "CREATE TABLE nf_behavior(
nf_behave_id integer NOT NULL PRIMARY KEY,
severity_id float,
flmlnA float,
flmlnE float,
flmlnH float,
flmlnL float,
flmlnM float,
ftype_E float,
ftype_H float,
ftype_L float,
ftype_M float,
ros float,
rosE float,
rosH float,
rosL float,
rosM float,
FOREIGN KEY(severity_id) REFERENCES severity(severity_id)
);")
```

Create non-forest risk table

```{r nf_risk, eval = FALSE, echo = TRUE}
dbExecute(term_project, "CREATE TABLE nf_risk(
nf_risk_id integer NOT NULL PRIMARY KEY,
severity_id float,
fri float,
fti float,
sdr varchar(12),
WHP float,
FOREIGN KEY(severity_id) REFERENCES severity(severity_id)
);")
```

Create non-forest_structure table

```{r nf_structure, eval = FALSE, echo = TRUE}
dbExecute(term_project, "CREATE TABLE nf_structure(
nf_str_id integer NOT NULL PRIMARY KEY,
severity_id float,
wwacbd float,
wwacbh float,
wwacc float,
wwach float,
wwafuel13 float,
wwafuel40 float,
FOREIGN KEY(severity_id) REFERENCES severity(severity_id)
);")
```

Create non-forest_topo table

```{r nf_topo, eval = FALSE, echo = TRUE}
dbExecute(term_project, "CREATE TABLE nf_topo(
nf_topo_id integer NOT NULL PRIMARY KEY,
severity_id float,
asp varchar(5),
dem float,
slope float,
FOREIGN KEY(severity_id) REFERENCES severity(severity_id)
);")
```


Create non-forest_veg table

```{r nf_veg, eval = FALSE, echo = TRUE}
dbExecute(term_project, "CREATE TABLE nf_veg(
nf_veg_id integer NOT NULL PRIMARY KEY,
severity_id float,
EVT varchar(60),
SystGrp varchar(60),
SAFSrm varchar(60),
NVCSOrd varchar(16),
NVCSclass varchar(25),
NVCSsubclass varchar(40),
SYSTMGRPPH varchar(30),
FOREIGN KEY(severity_id) REFERENCES severity(severity_id)
);")
```

