# Data visualization {#tidyverse}


```{r begin, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install packages `tidyverse` and `viridis`

```{r load_packages, eval = TRUE, echo = TRUE}
library(tidyverse)
library(viridis)
library(RSQLite)
library(DBI)
```

Connect to database 'severity' and create tables

```{r connection, eval=TRUE, echo=TRUE}
severity_db <- dbConnect(drv = RSQLite::SQLite(),
                "/Users/kiplingklimas/Box/Lab_Group/Kipling/Classes/WILD6900_EcoRepSci/Term_project/Term_project/database/severity.db")

severity <- dbGetQuery(severity_db, "SELECT * FROM severity;")
forest_behavior <- dbGetQuery(severity_db, "SELECT * FROM forest_behavior;")
forest_risk <- dbGetQuery(severity_db, "SELECT * FROM forest_risk;")
forest_structure <- dbGetQuery(severity_db, "SELECT * FROM forest_structure;")
forest_topo <- dbGetQuery(severity_db, "SELECT * FROM forest_topo;")
forest_veg <- dbGetQuery(severity_db, "SELECT * FROM forest_veg;")
nf_behavior <- dbGetQuery(severity_db, "SELECT * FROM nf_behavior;")
nf_risk <- dbGetQuery(severity_db, "SELECT * FROM nf_risk;")
nf_structure <- dbGetQuery(severity_db, "SELECT * FROM nf_structure;")
nf_topo <- dbGetQuery(severity_db, "SELECT * FROM nf_topo;")
nf_veg <- dbGetQuery(severity_db, "SELECT * FROM nf_veg;")
```

This is fire severity dataset shows the relationship between severity measurements and a variety of predictive variables
including topographic, vegetation, structural and derived risk indices. 
Data was differentiated between forest and non-forest points to compare severity relationships between forests and non-forest 
landcover systems. 

Simple scatter plot of forest severity by elevation (m)

```{r simple_plots, eval=TRUE, echo=TRUE}
forest_topo %>% 
  left_join(severity, by = "severity_id") %>% 
  select(dem, severity) %>% 
  ggplot(aes(x = dem, y = severity)) +
  geom_point() +
  labs(y = "Severity (dNBR)", x = "Elevation (m)") +
  theme_light()

```

Density curves of fire type extreme by severity

```{r fire_behavior, eval=TRUE, echo=TRUE}
forest_behavior %>% 
  left_join(severity, by = "severity_id") %>% 
  mutate(ftypeE = case_when(
    ftype_E == 3 ~ "Surface",
    ftype_E == 4 ~ "Passive",
    ftype_E == 5 ~ "Passive",
    ftype_E == 6 ~ "Active",
    ftype_E == 7 ~ "Active",
  )) %>% 
  ggplot(aes(x = severity, color = ftypeE)) +
  geom_density() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y = "Density", x = "Severity") 

```

Bar plot of canopy categories

```{r fbarplot, eval=TRUE, echo=TRUE}
forest_veg %>% 
  left_join(severity, by = "severity_id") %>% 
  filter(!is.na(severity_id)) %>% 
  ggplot(aes(x = NVCSsubclass, y = severity)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y = "Severity (dNBR)", x = "Vegetation type") 
```
