# Forest data visualization {#dataviz}


```{r f_begin, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install packages `tidyverse`, `viridis` and `ggridges`

```{r f load_packages, eval = TRUE, echo = FALSE}
library(tidyverse)
library(viridis)
library(RSQLite)
library(DBI)
library(ggridges)
```


```{r f_connection, eval=TRUE, echo=FALSE}
severity_db <- dbConnect(drv = RSQLite::SQLite(),
                "/Users/kiplingklimas/Box/Lab_Group/Kipling/Classes/WILD6900_EcoRepSci/Term_project/Term_project/database/severity.db")

severity <- dbGetQuery(severity_db, "SELECT * FROM severity;")
forest_behavior <- dbGetQuery(severity_db, "SELECT * FROM forest_behavior;")
forest_risk <- dbGetQuery(severity_db, "SELECT * FROM forest_risk;")
forest_structure <- dbGetQuery(severity_db, "SELECT * FROM forest_structure;")
forest_topo <- dbGetQuery(severity_db, "SELECT * FROM forest_topo;")
forest_veg <- dbGetQuery(severity_db, "SELECT * FROM forest_veg;")
```


```{r f_simple_plots, eval=TRUE, echo=TRUE}
sev <- forest_topo %>% 
  left_join(severity, by = "severity_id") %>% 
  select(dem, severity)

reg <- lm(formula= severity ~ dem, data = sev)
regs <- predict(reg, se.fit = TRUE)
regs <- data.frame(mean = regs$fit,
                    upr = regs$fit + 1.96 * regs$se.fit,
                    lwr = regs$fit - 1.96 * regs$se.fit)
            

forest_topo %>% 
  left_join(severity, by = "severity_id") %>% 
  select(dem, severity) %>% 
  ggplot(aes(x = dem, y = severity, color= dem)) +
  geom_point(alpha = 0.5) +
  scale_color_viridis_c(option = "magma") +
  geom_line(aes(y = regs$mean, color = regs$mean)) +
  labs(y = "Severity (dNBR)", x = "Elevation (m)") +
  theme_light() 
  

```

Simple scatter plot of forest severity vs. elevation (m)



```{r behavior, eval=TRUE, echo=TRUE}
forest_behavior %>% 
  left_join(severity, by = "severity_id") %>% 
  mutate(ftypeE = case_when(
    ftype_E == 3 ~ "Surface",
    ftype_E == 4 ~ "Passive",
    ftype_E == 5 ~ "Passive",
    ftype_E == 6 ~ "Active",
    ftype_E == 7 ~ "Active",
  )) %>% 
  ggplot(aes(y = ftypeE, x = severity, fill = ftypeE)) +
  ggridges::geom_density_ridges(scale = 8, alpha = 0.5, show.legend = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  +
  labs(y = "Fire type Extreme", x = "Severity") +
  scale_color_viridis_d(option = "magma") +
  theme_minimal()

```

Density curves of fire type extreme by severity
The variable FtypeE represents the fire type under extreme (99 percentile) weather scenarios. 
The three 'types' of fire in described by this variable are: 
1. Surface - fire that burns surface fuels 
2. Passive - fire that burns both surface and intermediate ladder fuels
3. Active - fire that burns in the canopy of the forest

This density plot is of interest because it shows that Active fire type represents lower severity burned areas than
Surface or Passive, which is contrary to a popular assumption that crown/canopy fire is associated with high-severity 
burned areas.

```{r ftypebarplot, eval=TRUE, echo=TRUE}
forest_veg %>% 
  left_join(severity, by = "severity_id") %>% 
  filter(!is.na(severity_id)) %>% 
  ggplot(aes(x = NVCSsubclass, y = severity)) +
  geom_boxplot() +
  labs(y = "Severity (dNBR)", x = "Canopy type") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

Bar plot of canopy categories

```{r risk_elevation, eval=TRUE, echo=TRUE}
forest_topo %>% 
  left_join(severity, by = "severity_id") %>% 
  left_join(forest_risk, by = "severity_id") %>% 
  select(dem, fri, severity) %>% 
  ggplot(aes(x = dem, y = severity, color = fri)) +
  geom_point(alpha = 0.5) +
  scale_color_viridis_c(option = "magma") +
  labs(y = "Severity (dNBR)", x = "Elevation (m)") +
  theme_light() 
  
```

Elevation x severity x fire risk index

```{r cc_elevation, eval=TRUE, echo=TRUE}
forest_topo %>% 
  left_join(severity, by = "severity_id") %>% 
  left_join(forest_structure, by = "severity_id") %>% 
  ggplot(aes(x = dem, y = severity, color = wwacc)) +
  geom_point(alpha = 0.5) +
  scale_color_viridis_c(option = "magma") +
  labs(y = "Severity (dNBR)", x = "Elevation (m)") +
  theme_light() 

```

Elevation x severity x canopy closure

```{r fuels_elevation, eval=TRUE, echo=TRUE}
forest_topo %>% 
  left_join(severity, by = "severity_id") %>% 
  left_join(forest_structure, by = "severity_id") %>% 
  ggplot(aes(x = dem, y = severity, color = wwafuel40)) +
  geom_point(alpha = 0.5) +
  scale_color_viridis_c(option = "magma") +
  labs(y = "Severity (dNBR)", x = "Elevation (m)") +
  theme_light()

```

Elevation x severity x fuels40

Mean severity by vegetation type

``` {r f_vegtype, eval = TRUE, echo = TRUE}
forest_veg %>% 
  pull(EVT) %>% 
  unique()
```

Lets check out the mean severity for each of these forest types. 
The red dashed line is the high-severity threshold. 
There are three forest types that have an average severity at or above the high-severity threshold of about 450 dNBR.
The three forest types are: 

1. Rocky Mountain Subalpine Dry-Mesic Spruce-Fir Forest and Woodland
2. Rocky Mountain Subalpine-Montane Limber-Bristlecone Pine Woodland
3. Southern Rocky Mountain Mesic Montane Mixed Conifer Forest and Woodland


```{r vegtype_mean_sev, eval = TRUE, echo = TRUE}
forest_veg %>% 
  left_join(severity, by = "severity_id") %>% 
  group_by(EVT) %>% 
  summarise(mean_severity = mean(severity, na.rm = TRUE)) %>% 
  ggplot(aes(x = EVT, y = mean_severity)) +
  geom_hline(yintercept = 450, linetype = "dashed", color = "red") +
  geom_bar(stat = "identity") +
  labs(x = " ", y = "Mean severity") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 40, hjust = 1)) 
  
```