# Non-forest data visualization {#ggplot}


```{r nf_begin, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages 

```{r nf_load_packages, eval = TRUE, echo = FALSE}
library(tidyverse)
library(viridis)
library(RSQLite)
library(DBI)
library(ggridges)
```


```{r nf_connection, eval=TRUE, echo=FALSE}
severity_db <- dbConnect(drv = RSQLite::SQLite(),
                "/Users/kiplingklimas/Box/Lab_Group/Kipling/Classes/WILD6900_EcoRepSci/Term_project/Term_project/database/severity.db")

severity <- dbGetQuery(severity_db, "SELECT * FROM severity;")
nf_behavior <- dbGetQuery(severity_db, "SELECT * FROM nf_behavior;")
nf_risk <- dbGetQuery(severity_db, "SELECT * FROM nf_risk;")
nf_structure <- dbGetQuery(severity_db, "SELECT * FROM nf_structure;")
nf_topo <- dbGetQuery(severity_db, "SELECT * FROM nf_topo;")
nf_veg <- dbGetQuery(severity_db, "SELECT * FROM nf_veg;")
```


Simple scatter plot of non-forest severity vs. elevation (m)


```{r nf_scatterplot, eval=TRUE, echo=TRUE}
sev <- nf_topo %>% 
  left_join(severity, by = "severity_id") %>% 
  select(dem, severity)

reg <- lm(formula= severity ~ dem, data = sev)
regs <- predict(reg, se.fit = TRUE)
regs <- data.frame(mean = regs$fit,
                    upr = regs$fit + 1.96 * regs$se.fit,
                    lwr = regs$fit - 1.96 * regs$se.fit)
            

nf_topo %>% 
  left_join(severity, by = "severity_id") %>% 
  select(dem, severity) %>% 
  ggplot(aes(x = dem, y = severity, color= dem)) +
  geom_point(alpha = 0.5) +
  scale_color_viridis_c(option = "magma") +
  geom_line(aes(y = regs$mean, color = regs$mean)) +
  labs(y = "Severity (dNBR)", x = "Elevation (m)") +
  theme_light() 
  

```


```{r nf_barplot, eval=TRUE, echo=TRUE}
nf_veg %>% 
  left_join(severity, by = "severity_id") %>% 
  filter(!is.na(severity_id)) %>% 
  ggplot(aes(x = NVCSsubclass, y = severity)) +
  geom_boxplot() +
  labs(y = "Severity (dNBR)", x = "Canopy type") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

Bar plot of non-forest canopy categories.

```{r nf_risk_elevation, eval=TRUE, echo=TRUE}
nf_topo %>% 
  left_join(severity, by = "severity_id") %>% 
  left_join(nf_risk, by = "severity_id") %>% 
  select(dem, fri, severity) %>% 
  ggplot(aes(x = dem, y = severity, color = fri)) +
  geom_point(alpha = 0.5) +
  scale_color_viridis_c(option = "magma") +
  labs(y = "Severity (dNBR)", x = "Elevation (m)") +
  theme_light() 
  
```

Elevation x severity x fire risk index

```{r nf_cc_elevation, eval=TRUE, echo=TRUE}
nf_topo %>% 
  left_join(severity, by = "severity_id") %>% 
  left_join(nf_structure, by = "severity_id") %>% 
  ggplot(aes(x = dem, y = severity, color = wwacc)) +
  geom_point(alpha = 0.5) +
  scale_color_viridis_c(option = "magma") +
  labs(y = "Severity (dNBR)", x = "Elevation (m)") +
  theme_light() 

```

Elevation x severity x canopy closure. This figures illustrates that non-forest landcover has no discernable canopy. 

```{r nf_fuels_elevation, eval=TRUE, echo=TRUE}
nf_topo %>% 
  left_join(severity, by = "severity_id") %>% 
  left_join(nf_structure, by = "severity_id") %>% 
  ggplot(aes(x = dem, y = severity, color = wwafuel40)) +
  geom_point(alpha = 0.5) +
  scale_color_viridis_c(option = "magma") +
  labs(y = "Severity (dNBR)", x = "Elevation (m)") +
  theme_light()

```

Elevation x severity x fuels40

Mean severity by vegetation type

``` {r nf_vegtype, eval = TRUE, echo = TRUE}
nf_veg %>% 
  pull(EVT) %>% 
  unique()
```

Lets check out the mean severity for each of these forest types. 
The red dashed line is the high-severity threshold. 

```{r nf_vegtype_mean_sev, eval = TRUE, echo = TRUE}
nf_veg %>% 
  left_join(severity, by = "severity_id") %>% 
  group_by(EVT) %>% 
  summarise(mean_severity = mean(severity, na.rm = TRUE)) %>% 
  ggplot(aes(x = EVT, y = mean_severity)) +
  geom_hline(yintercept = 450, linetype = "dashed", color = "red") +
  geom_bar(stat = "identity") +
  labs(x = " ", y = "Mean severity") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 40, hjust = 1)) 
  
```